---
title: "ANZ Virtual Internship EDA"
author: "Gunay Mammadsalahova"
date: "8/20/2020"
output: html_document
---

The dataset contains 12043 transactions for 100 customers who have one bank account each.Trasactional period is from 01/08/2018 - 31/10/2018 (92 days duration).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## EXPLORATORY DATA ANALYSIS


```{r}
anz <- read.csv("anz.csv", na.strings = c("NA",""))
dim(anz)
```
Data have 12043 rows and 23 columns.
```{r}
#columns
colnames(anz)
```

```{r}
library(tidyverse)
glimpse(anz)
```

```{r}
#Columns with missing values
colSums(is.na(anz))
```
```{r}
#delete columns with 92% NA's
anz <- anz[,-c(3,9)]
```
So, we deleted `bpay_biller_code` and `merchant_code` columns.
```{r}
library(Amelia)
missmap(anz, col = c("black","grey"))
```
As we see, missing values in columns overlap.
```{r}
#Average transaction amount
mean(anz$amount) # 187.9336
```
So, the average of the `amount` is `r mean(anz$amount)`.

```{r}
#gender 
library(ggplot2)
ggplot(anz, aes(x = gender))+
  geom_bar(fill = "purple") +
  labs(x = "Gender", y = "Transaction count",
       title = "Transaction counts for each gender")
```
```{r echo=FALSE}
table(anz$gender)
```

```{r echo=FALSE}
#movement
ggplot(anz, aes(x = movement))+
  geom_bar(fill = "light blue") +
  labs(x = "Transaction Movement", y = "Number of Transactions",
       title = "Transaction Movement")
```
```{r}
table(anz$movement)
```
```{r echo=FALSE}
#description
ggplot(anz, aes(x = txn_description))+
  geom_bar(fill = "yellow") +
  labs(x = "Description", y = "Number of Transactions",
       title = "Transaction description")
```

```{r echo=FALSE}
table(anz$txn_description)
```
```{r echo=FALSE}
#merchant state
ggplot(anz, aes(x = merchant_state))+
  geom_bar(fill = "blue")+
  labs(x = "Merchant state",y = "Number of Transactions",
       title = "Transactions in each merchant state")
```
```{r echo=FALSE}
table(anz$merchant_state)
```

```{r echo=FALSE}
##historgram of total amount transacted
mean_trans<-anz %>%
  group_by(customer_id) %>%
  summarise(average_trans=n()/3)

ggplot(mean_trans,aes(average_trans))+geom_histogram(bins = 20, fill = "salmon")+
  labs(x="Monthly Average Transactions",y="Number of Customers",title="Monthly Average Transactions")
```
Maximum transaction amount is around 200 AUD
```{r}
#changing date column's format
anz$date<-as.Date(anz$date,format = "%m/%d/%Y")
##extracting weekday of transaction
anz$weekday<-weekdays(anz$date)
```

```{r}
library(lubridate)
#extracting time of transaction
anz$hour<-hour(as.POSIXct(substr(anz$extraction,12,19),format="%H:%M:%S"))
```

```{r echo=TRUE}
#Mean of spending during one week
mean_spending_aug_week<- anz %>% filter(movement=="debit")%>%filter(date >="2018-08-06" & date<="2018-08-12") %>% group_by(date) %>% summarise(mean_spending=mean(amount))
ggplot(mean_spending_aug_week,aes(x=date,y=mean_spending))+geom_line() +
  labs(x = "Date", y = "Mean of spending $", title = "Mean of spending during one week")
```
So,here the maximum average of spending is occured in Tuesday, the least in Friday.

```{r echo=TRUE}
#Average volume of transactions during one week
mean_volume_aug_week<- anz %>%filter(date >="2018-08-06" & date<="2018-08-12") %>% group_by(date) %>% summarise(average_volume=n())
ggplot(mean_volume_aug_week,aes(x=date,y=average_volume))+geom_line() + 
  labs(x = "Date", y = "Average volume",title = "Average volume of transactions during one week")

```
The least average volume of transactions happened in Monday, the most in Friday.

```{r  echo=TRUE}
## Visualise transaction volume over an average week.
atw <- anz %>% select(date, weekday) %>% group_by(date, weekday) %>%
  summarise(daily_avg_vol = n()) %>% group_by(weekday) %>%
  summarise(avg_vol=mean(daily_avg_vol, na.rm = TRUE))
atw$weekday <- factor(atw$weekday, levels = c("Monday","Tuesday","Wednesday",
                                              "Thursday","Friday","Saturday","Sunday"))
ggplot(atw, aes(x = weekday, y = avg_vol)) + geom_point() + geom_line(aes(group = 1))+
  ggtitle('Average transaction volume by weekday')+
  labs(x ='Weekday',y='Transaction volume')

```
```{r echo=TRUE}
# visualize transaction volume over an average week.
ath <- anz %>% select(date,hour) %>%  group_by(date,hour) %>% summarize(trans_vol=n()) %>%   group_by(hour) %>% summarize(trans_vol_per_hr = mean(trans_vol,na.rm=TRUE))
ggplot(ath,aes(x=hour,y=trans_vol_per_hr))+geom_point()+geom_line(aes(group = 1))+  ggtitle('Average transaction volume by hour') +  labs(x='Hour',y='Transaction volume') + expand_limits( y = 0)

```

